---
title: "Segmnetation and Clustering"
author: "Bhoomika John Pedely"
date: "2023-09-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a mini-project inspired by the 'Business Decision Algorithm' class during my MSBA program. I will be using a unsupervised clustering algorithm to create customer segments of 50 customers, based on 2 variables, 'deal proneness' and 'frequency of purchase'.

```{r}
#Loading the data
library(gplots)
library(RColorBrewer)
library(dendextend)
library(circlize)
library(NbClust)

dataC <- read.csv("http://data.mishra.us/files/dataClust.csv")

#scatter plot for exploration
plot(dataC$deal, dataC$frequency, xlab="Deal proneness",
ylab="Frequency of purchase")
```
The scatterplot doesn't show any linear relationship between the two variables of interest. Next, I will be looking at dendograms made using 4 different types of linkages. 

```{r}
d <- dist(dataC, method = "euclidean")

fit.w <- hclust(d,method="ward.D2")
plot(fit.w, main = "Ward Linkage", xlab = "Observations", ylab = "Height", cex = .6)
```
```{r}
fit.c <- hclust(d,method="complete")

plot(fit.c, main = "CompleteLinkage", xlab = "Observations", ylab = "Height", cex = .6)
13
```
```{r}
fit.s <- hclust(d,method="single")

plot(fit.s, main = "Single Linkage", xlab = "Observations", ylab = "Height", cex = .6)
```
```{r}
fit.a <- hclust(d,method="average")

plot(fit.a, main = "Average Linkage", xlab = "Observations", ylab = "Height", cex = .6)
```
Evaluating Cophenetic correlation coefficient for all the types of linkages. 
```{r}
(d.coph <- cophenetic(fit.w)) # ward linkage
cor(d, d.coph) # 0.6645585
(d.coph <- cophenetic(fit.c)) # complete linkage
cor(d, d.coph) # 0.65408
(d.coph <- cophenetic(fit.s)) # single linkage
cor(d, d.coph) # 0.444397
(d.coph <- cophenetic(fit.a)) # average linkage
cor(d, d.coph) #0.741388
```
```{r}
cutree(fit.c, 3)
```
```{r}
cutree(fit.s, 3)
```
```{r}
cutree(fit.w, 3)
```
How many clusters to segment customers into?
```{r}
res<-NbClust(dataC, distance = "euclidean", min.nc=2, max.nc=8,
method = "ward.D2", index = "all")
```
The Hubert index is a graphical method of determining the number of clusters. In the plot of Hubert index, we seek a significant knee that corresponds to a significant increase of the value of the measure i.e the significant peak in index second differences plot.

The D index is a graphical method of determining the number of clusters. In the plot of D index, we seek a significant knee (the significant peak in Dindex second differences plot) that corresponds to a significant increase of the value the measure.

Heatmap

Assuming that deal proneness and frequency of visit convey information about customers that can be used by the store, what type of insights can
be drawn? 

Would we expect deal prone customers to visit the store more frequently or
would we expect to see the opposite pattern of results? 

Does a customer who has a high value (green color) on deal proneness also have a high value of frequency of visits? If yes, then what can we infer about these customers?

```{r}
my_palette <- colorRampPalette(c('red','yellow','green'))(256)
heatmap.2(as.matrix(dataC),
cexRow=0.5,
cexCol=1.1,
col = my_palette,
sepcolor="black",
sepwidth=c(0.01,0.01),
scale="none",
trace="none")
```

